---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# deduped

<!-- badges: start -->
[![](https://cranlogs.r-pkg.org/badges/deduped)](https://cran.r-project.org/package=deduped)
<!-- badges: end -->

`deduped` contains one main function `deduped()` which speeds up slow,
vectorized functions by only performing computations on the unique values
of the input and expanding the results at the end. A convenience wrapper,
`with_deduped()`, was added in version 0.3.0 to allow piping an existing 
expression.

## Installation

You can install the released version of `deduped` from
[CRAN](https://cran.r-project.org/package=deduped) with:

``` r
install.packages("deduped")
```

And the development version from [GitHub](https://github.com/orgadish/deduped):

``` r
if(!requireNamespace("remotes")) install.packages("remotes")

remotes::install_github("orgadish/deduped")
```

## Examples

### Setup
```{r}
library(deduped)
set.seed(0)

slow_tolower <- function(x) {
  for (i in x) {
    Sys.sleep(0.0005)
  }
  tolower(x)
}
```


### `deduped()`

```{r example}

# Create a vector with significant duplication.
unique_vec <- sample(LETTERS, 5)
duplicated_vec <- sample(rep(unique_vec, 50))
length(duplicated_vec)

system.time({  x1 <- slow_tolower(duplicated_vec)  })


system.time({  x2 <- deduped(slow_tolower)(duplicated_vec)  })
all.equal(x1, x2)
```
_Note: As of version 0.3.0, you could also use_ `slow_tolower(duplicated_vec) |> with_deduped()`.

### `deduped(lapply)()`

`deduped()` can also be combined with `lapply()` or `purrr::map()`.

```{r example-map}

unique_list <- lapply(1:3, function(j) sample(LETTERS, j, replace = TRUE))
str(unique_list)

# Create a list with significant duplication.
duplicated_list <- sample(rep(unique_list, 50)) 
length(duplicated_list)

system.time({  y1 <- lapply(duplicated_list, slow_tolower)  })
system.time({  y2 <- deduped(lapply)(duplicated_list, slow_tolower)  })

all.equal(y1, y2)
```
