---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# deduped

<!-- badges: start -->
[![](https://cranlogs.r-pkg.org/badges/deduped)](https://cran.r-project.org/package=deduped)
<!-- badges: end -->

`deduped` contains one main function `deduped()` which speeds up slow,
vectorized functions by only performing computations on the unique values
of the input and expanding the results at the end. A convenience wrapper,
`with_deduped()`, was added in version 0.3.0 to allow piping an existing 
expression.

Note: It only works on functions that preserve length and order.

## Installation

You can install the released version of `deduped` from
[CRAN](https://cran.r-project.org/package=deduped) with:

``` r
install.packages("deduped")
```

And the development version from [GitHub](https://github.com/orgadish/deduped):

``` r
if(!requireNamespace("remotes")) install.packages("remotes")

remotes::install_github("orgadish/deduped")
```

## Examples

### Setup
```{r}
library(deduped)
set.seed(0)

slow_tolower <- function(x) {
  for (i in x) {
    Sys.sleep(0.0005)
  }
  tolower(x)
}
```


### `deduped(...)`

```{r example}

# Create a vector with significant duplication.
set.seed(1)
unique_vec <- sample(LETTERS, 5)
print(unique_vec)
duplicated_vec <- sample(rep(unique_vec, 100))
length(duplicated_vec)

system.time({  x1 <- slow_tolower(duplicated_vec)  })


system.time({  x2 <- deduped(slow_tolower)(duplicated_vec)  })
all.equal(x1, x2)

# As of version 0.3.0, you can use `with_deduped()`.
all.equal(x1, slow_tolower(duplicated_vec) |> with_deduped())
```


### `deduped(lapply)(...)`

`deduped()` can also be combined with `lapply()` or `purrr::map()`.

```{r example-lapply}

set.seed(2)
unique_list <- lapply(1:3, function(j) sample(LETTERS, j, replace = TRUE))
str(unique_list)

# Create a list with significant duplication.
duplicated_list <- sample(rep(unique_list, 50)) 
length(duplicated_list)

system.time({  y1 <- lapply(duplicated_list, slow_tolower)  })
system.time({  y2 <- deduped(lapply)(duplicated_list, slow_tolower)  })

all.equal(y1, y2)
```

### `deduped(fs::path_rel)(...)`

`deduped()` is helpful on slow path functions like `fs::path_rel()`.

```{r example-path_rel}

set.seed(3)
top_path <- "x/y/z/"
unique_paths <- paste0(top_path, LETTERS, "/file.csv")
str(unique_paths)

# Create a list with significant duplication.
dup_paths <- sample(rep(unique_paths, 500)) 
length(dup_paths)

system.time({  y1 <- fs::path_rel(dup_paths, start=top_path)  })
system.time({  y2 <- deduped(fs::path_rel)(dup_paths, start=top_path)  })

all.equal(y1, y2)
```

### When to use `with_deduped()`

`with_deduped()` is a convenience wrapper for interactive or one-off use.
Because it reconstructs the wrapper function on every call, prefer `deduped()`
directly when calling inside a loop:

```r
# Good: wrapper is built once
deduped_slow_tolower <- deduped(slow_tolower)
for (x in list_of_vecs) deduped_slow_tolower(x)

# Avoid: wrapper is rebuilt on every iteration
for (x in list_of_vecs) slow_tolower(x) |> with_deduped()
```
